---
title: "Spatiotemporal transmission risk"
author: "Nicolas Banholzer"
date: "2022-11-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform = FALSE)
```

## Libraries

```{r}
library(tidyverse)
library(reshape2)
library(rriskDistributions)
library(tidybayes)
library(lubridate)
library(sf)
library(raster)
library(terra)
library(ggridges)
library(grid)
library(ggplot2)
library(ggrepel)

source("../utils/distr.r")
source("../utils/trans_risk.r")
source("../utils/spatial.r")
source("../utils/plotting.R")

# plot with and height of input and data graphs
width_descr <- 10.5 / cm(1)
height_descr <- 7.5 / cm(1)
text_descr <- 12
```


## Data

```{r}
selected_dates <- as.Date(list.files("../data-clean/Masi/patient-tracking-data")[-1])
```

### Clinical

```{r}
clinical_dta <- readRDS("../data-clean/Masi/clinical-data/clinical-data.rds") %>%
  filter(date %in% selected_dates)

# number of registered patients
clinical_dta %>%
  group_by(patient_id) %>%
  slice(1) %>%
  nrow()

# share of TB patients
share_tb <- clinical_dta  %>%
  group_by(date, patient_id) %>%
  summarize(tb = any(tb_suspect == "yes")) %>%
  ungroup() %>%
  group_by(date) %>%
  summarize(n = n(),
            n_tb = sum(tb)) %>%
  arrange(date) %>%
  mutate(prop_tb = n_tb / n * 100,
         date = format(date, format = "%b %d"),
         date = factor(date, levels = unique(date))) %>%
  ungroup() 
share_tb_pl <- share_tb %>%
  ggplot(aes(x = date, y = n_tb)) +
  geom_bar(stat = "identity", fill = "grey", color = "black", width = .5) +
  annotate("text", x = "Nov 10", y = mean(share_tb$n_tb), label = paste0("Mean = ",round(mean(share_tb$n_tb), 1)), color = "darkred", size = 8 / cm(1), vjust = -1) +
  geom_hline(aes(yintercept = mean(share_tb$n_tb)), color = "darkred") +
  scale_y_continuous(expand = expansion(mult = c(0,0.05))) +
  scale_x_discrete(breaks = format(selected_dates, format = "%b %d")[c(1, 4, 8, 12)]) +
  labs(y = "No. of masked TB patients") +
  theme_classic() +
  theme(axis.title.x = element_blank(),
        plot.margin = margin(5.5, 5.5, 5.5, 5.5),
        text = element_text(size = text_descr))
ggsave(plot = share_tb_pl, filename = "../doc/IWHOD/presentation/figures/share-tb-patients.png", w = width_descr, h = height_descr)
```

### CO2

```{r}
co2_pl <- readRDS("../data-clean/Masi/environmental-data/co2-temp-humid.rds") %>%
  filter(location %in% c("waiting room", "tb room"),
         date %in% selected_dates) %>%
  mutate(location = factor(ifelse(location == "tb room", "TB room", "Waiting room"), levels = c("Waiting room", "TB room")),
         time = as.POSIXct(time, format = "%H:%M", tz = "CET")) %>%
  ggplot(aes(x = time, y = co2, color = location, fill = location)) +
  stat_lineribbon(alpha = .25, size = .75) +
  scale_x_datetime(expand = c(0,0), date_breaks = "2 hours", date_labels = "%H:%M") +
  scale_y_continuous(expand = c(0,0), breaks = scales::breaks_width(width = 100)) +
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  scale_color_manual(values = wes_palette("Moonrise2")) +
  labs(y = expression("CO"[2]*" (ppm)"), fill = "Quantile", color = "Quantile") +
  theme_classic() +
  theme(legend.position = c(.65,.95), legend.direction = "horizontal", legend.key.width = unit(0.3, "cm"), legend.key.height = unit(0.2, "cm"),
        plot.margin = margin(5.5, 5.5, 5.5, 5.5), axis.title.x = element_blank(), legend.title = element_blank(),
        text = element_text(size = text_descr)) 
ggsave(plot = co2_pl, filename = "../doc/IWHOD/presentation/figures/co2-levels.png", width = width_descr, height = height_descr)
```

### Building

```{r}
# background image
clinic_img <- png::readPNG("../data-raw/Masi/building/clinic_clipped.png")

# building
cellSize <- sqrt(1 * 1e9 / 3e3) 
building <- terra::vect(paste("../data-raw", "Masi", "building", "clinic-vector.gpkg", sep = "/"))
building_sf <- sf::st_as_sf(building, crs = NA)
sf::st_crs(building_sf) <- NA
waiting_room <- shapeToSpatial(building_sf$geometry[2], cellSize)
waiting_room_df <- fortify(waiting_room)
passage <- shapeToSpatial(building_sf$geometry[4], cellSize)
passage_df <- fortify(passage)
tb_room <- shapeToSpatial(building_sf$geometry[3], cellSize)
tb_room_df <- fortify(tb_room)

# test
waiting_room_df$value <- rnorm(nrow(waiting_room_df))
passage_df$value <- rnorm(nrow(passage_df))
tb_room_df$value <- rnorm(nrow(tb_room_df))

ggplot(mapping = aes(x = long, y = lat, group = group, fill = value)) +
  geom_polygon(data = waiting_room_df) +
  geom_polygon(data = passage_df) +
  geom_polygon(data = tb_room_df) +
  ggpubr::background_image(clinic_img) +
  annotate("text", x = -13200, y = -5300, label = "Entrance", size = 8 / cm(1), fontface = 2, hjust=0,vjust=0) +
  annotate("text", x = -13400, y = -2300, label = "Waiting room", size = 8 / cm(1), fontface = 2, hjust=0,vjust=0) +
  annotate("text", x = -5000, y = -4100, label = "Registration", size = 8 / cm(1), fontface = 2, angle = 90, hjust=0,vjust=0) +
  annotate("text", x = 2600, y = -2700, label = "TB room", size = 8 / cm(1), fontface = 2) +
  scale_y_continuous(expand = expansion(add=c(2500, 4000))) +
  scale_x_continuous(expand = expansion(add=c(400, 500))) +
  labs(fill = "Density") +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = c(0.825,0.075), legend.direction = "horizontal", legend.title = element_blank(), #legend.title = element_text(vjust = .75),
        panel.background = element_rect(fill = RColorBrewer::brewer.pal(n = 3, name = "YlOrRd")[1]), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```


### No. people

```{r}
files <- list.files("../data-clean/Masi/patient-tracking-data/", full.names = T)[-1]
no_people <- do.call(rbind, lapply(files, function(x) readRDS(paste0(x, "/patient-id-data_matched-filtered.rds")))) %>%
  mutate(date = as.Date(format(time, format = "%Y-%m-%d")),
         time_min = as.POSIXct(format(time, format = "%H:%M"), format = "%H:%M"))

# number of clinical attendees
no_people %>%
  group_by(date, patient_id) %>%
  slice(1) %>%
  nrow()
  
# over time
no_people_time <- no_people %>%
  group_by(date, time_min, time) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  group_by(date, time_min) %>%
  summarize(n = mean(n)) %>%
  ungroup()

no_people_time_pl <- no_people_time %>%
  ggplot(aes(x = time_min, y = n)) +
  stat_lineribbon(alpha = .25, size = .5) +
  scale_x_datetime(expand = c(0,0), date_breaks = "2 hours", date_labels = "%H:%M") +
  scale_y_continuous(expand = c(0,0), limits = c(0, NA), breaks = scales::breaks_width(5)) +
  scale_fill_brewer(palette = "OrRd") +
  labs(y = "No. of clinical attendees", fill = "Quantile", color = "Quantile") +
  theme_bw2() +
  theme(axis.title.x = element_blank(), legend.position = c(.8,.9), legend.direction = "horizontal", legend.key.width = unit(0.3, "cm"), legend.key.height = unit(0.2, "cm"),
        plot.margin = margin(5.5, 10, 5.5, 5.5)) +
  guides(fill = guide_legend(title.position = "top"))
no_people_time_pl

# Spatial
no_people_spatial_pl <- no_people %>%
  ggplot(aes(x = x, y = y)) +
  geom_bin2d(binwidth = 100) +
  scale_fill_stepsn(colours = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd"), trans = "log", breaks = c(1, 10, 100, 1000, 10000)) +
    ggpubr::background_image(clinic_img) +
    annotate("text", x = -13200, y = -5800, label = "Entrance", size = text_descr / cm(1), hjust=0,vjust=0) +
    annotate("text", x = -14000, y = -2600, label = "Waiting room", size = text_descr / cm(1), hjust=0,vjust=0) +
    annotate("text", x = -5800, y = -4500, label = "Registration", size = text_descr / cm(1), angle = 90, hjust=0,vjust=0) +
    annotate("text", x = 3000, y = -2700, label = "TB room", size = text_descr / cm(1)) +
    scale_y_continuous(expand = expansion(mult=c(0.05, 0.15))) +
    scale_x_continuous(expand = expansion(mult=c(0.03, 0))) +
    labs(fill = "Density") +
    theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.key.width = unit(0.3, "cm"), legend.key.height = unit(0.3, "cm"),
          legend.position = c(0.825,0.075), legend.direction = "horizontal",legend.title = element_text(vjust = .75),
          panel.background = element_rect(fill = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd")[1]), panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()) +
  guides(fill = guide_legend(title.position = "top", label.position = "bottom"))
ggsave(plot=no_people_spatial_pl, filename = "../doc/IWHOD/presentation/figures/no-people-spatial.png", width = width_descr, height = height_descr)
```


## Assumptions

```{r}
n_samples <- 1000
set.seed(12345)
```

### Spatial emission

* Wang (2021): 
  * Travel distance of particles depends on particle size, activity, and environmental factors
  * Smaller aerosols can travel further, stay in the air for a longer time, and typically contain higher viral load
  * The concentration of exhaled aerosols is highest close to the source and decreases with distance
  * Short-range aerosol transmission difficult to prevent with ventilation
* Chen (2020): Short-range aerosols dominate over most distance up to 2m, droplets only play a role at very close distance <0.5
* Ko (2004): exponential decrease of potential infections on an airplance (high ventilation) with distance to infector


--> Assume an exponential decrease in viral concentration with distance to source relative to the source (relative concentration = 1). 


```{r}
relQC_pl <- expand.grid(distance = seq(0, 10, .1), 
                           mu = rMuIc(n_samples)) %>%
  mutate(relQC = dIc(distance, mu) * 100) %>%
  ggplot(aes(x = distance, y = relQC)) +
  stat_lineribbon(alpha = .5, size = .5) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_brewer(palette = "OrRd") +
  labs(x = "Distance (m)", y = "Relative quanta concentration (%)", fill = "Quantile") +
  theme_bw2() +
  theme(legend.position = c(.8,.9), legend.direction = "horizontal", legend.key.width = unit(0.3, "cm"), legend.key.height = unit(0.2, "cm"),
        plot.margin = margin(5.5, 10, 5.5, 5.5)) +
  guides(fill = guide_legend(title.position = "top"))
save_plot(relQC_pl, pdf_file = "../results/inputs/spatial-relative-quanta-concentration.pdf", w = aw, h = ah)
```


### Infectious quanta 

* Riley (1962): 130 patients, q: 1.25
* Escombe (2008): 117 patients, q: 8.2; but high values above 100 are possible
* Nardell (1991) : 1 patients, q: 12.5
* Andrews (2014) : 571 patients, q: 0.89

--> Take the mean and standard deviation of these values and use a student-t distribution with small degrees of freedom to consider outliers

```{r}
q <- c(1.25, 8.2, 12.5, 0.89)
n_q <- c(130, 117, 1, 571)
mean_q <- sum(n_q * q) / sum(n_q)
sd_q <- sqrt( sum( (n_q / sum(n_q) * (q - mean_q) ^ 2 ) ) )

qLiterature <- data.frame(q = q, 
                          y = .01, 
                          lab = c("Riley", "Escombe", "Nardell", "Andrews"))

q_pl <- data.frame(q = seq(0, 40, .1)) %>%
  mutate(density = dq(q)) %>%
  ggplot(aes(x = q, y = density)) +
  geom_line() +
  geom_vline(mapping = aes(xintercept = 2), linetype = "dashed", color = "darkred") +
  annotate("text", x = 3, y = 0.175, hjust = 0, label = "Mean = 2.0", size = 8 / cm(1), color = "darkred") +
  geom_point(data = qLiterature, mapping = aes(x = q, y = y, color = lab, shape = lab)) +
  geom_text_repel(data = qLiterature, mapping = aes(x = q, y = y, color = lab, label = lab), size = 8 / cm(1)) +
  annotate("text", x = 30, y = 0.175, label = expression("q ~ Student-t("*mu*"=2.0,"*sigma*"=2.5"*","*nu*"=1)"), size = 8 / cm(1)) +
  scale_x_continuous(expand = c(0,0), limits = c(0, 40)) +
  scale_y_continuous(expand = expansion(mult = c(0,0.05))) +
  scale_color_brewer(palette = "Dark2", direction = -1) +
  labs(x = "Quanta emission (log q/h)", y = "Density") +
  theme_bw2() +
  theme(plot.margin = margin(5.5, 10, 5.5, 5.5), legend.position = "none")
save_plot(q_pl, pdf_file = "../results/inputs/quanta-emission.pdf", w = aw, h = ah)
```


### Unmasked TB patients

* Berhanu (2023): 
  * study group: clinic attendees living with HIV, contact with TB patients in past year, or had TB in the past 2 years
  * finding: 55% of  with a sputum result positive for M. tuberculosis did not have a positive TB symptom screen
* Hamada (2018): The WHO 4-question symptom screen (cough, fever, weight loss, and night sweats) misses up to half the TB cases among people living with human immunodeficiency virus (HIV) on antiretroviral therapy (ART)
--> Unmasked TB patients: Poisson(mu = No. of TB suspects)

```{r}
unmaskedTB <- data.frame(n_tb = share_tb$n_tb) %>%
  mutate(unmasked_tb = lapply(n_tb, function(x) rTBunmasked(n = n_samples, lambda = x))) %>%
  unnest(unmasked_tb) %>%
  group_by(unmasked_tb) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(prop_unmasked_tb = n / sum(n)) 
mean_unmaskedTB <- sum(unmaskedTB$unmasked_tb * unmaskedTB$prop_unmasked_tb)
unmaskedTB_pl <- unmaskedTB %>%
  ggplot(aes(y = prop_unmasked_tb, x = unmasked_tb)) +
  geom_bar(stat="identity", fill = "grey", color = "black") +
  geom_vline(mapping = aes(xintercept = mean_unmaskedTB), linetype = "dashed", color = "darkred") +
  annotate("text", y = 0.25, x = mean_unmaskedTB, hjust = -.2, label = paste0("Mean = ", round(mean_unmaskedTB, 1)), size = 8 / cm(1), color = "darkred") +
  scale_y_continuous(expand = expansion(mult = c(0,0.05))) +
  scale_x_continuous(expand = c(0,0)) +
  labs(x = "No. of unmasked TB patients", y = "Probability (%)") +
  theme_bw2()
save_plot(unmaskedTB_pl, pdf_file = "../results/inputs/unmasked-tb-patients.pdf", w = aw, h = ah)
```


### Ventilation 

Average volume of exhaled gas:
* Richardson et al. (2014): 0.1 - 0.17 l/s per person based on Feher (2017)
--> Use Uniform(0.1, 0.17)

CO2 levels in outdoor air:
* https://gml.noaa.gov/dv/data/index.php?site=CPT&parameter_name=Carbon%2BDioxide,%20accessed%2012%20November%202019
--> Set to 4125.5

CO2 levels in exhaled air:
* Rudnick and Milton (2003): 40,000
--> Set to 40,000

CO2 generation rate
* Persily et al.: see Table 5, around 0.003 to 0.005
--> Use Uniform(0.003, 0.005)



### Inactivation

* Nardell (2016) reviews Loudon et al. (1969) and Lever et al. (2000) reporting contradictory findings of half-lives between 20m to 6h. Exact viability in the air is unknown.
* The finding by Lever et al. would suggest that sustained contact is required for transmission.
* The viability in the air probably depends a lot on the conditions, in particular ventilation and UV.  

--> TODO: Check how high the ventilation rate is compared to half-lives as to whether we can ignore the half-lives.


### Deposition onto surfaces

* Wang et al. (2021): The larger the size of the particle, the less long it will reside in the air. Environmental factors such as temperature, humidity and ventilation also play an important role.


## Evaluation

```{r}
sim_results <- list.files("../simulations/Masi/2021-10-25/", full.names = T)
sim_results <- do.call(rbind, lapply(sim_results, readRDS))
```

### Transmission risk

```{r}
# individual patient transmission risk by simulation
transRisk <- sim_results %>%
  group_by(patient_id, sim) %>%
  slice(1) %>%
  ungroup()

# Average transmission risk across patients
transRisk_av <- transRisk %>%
  group_by(sim) %>%
  summarize(across(c(P_N, P_N_ns), mean)) %>%
  ungroup()

mean_risk <- round(mean(transRisk_av$P_N * 100), 1)
q25_risk <- round(quantile(transRisk_av$P_N * 100, .25), 1)
q75_risk <- round(quantile(transRisk_av$P_N * 100, .75), 1)
message(sprintf("Spatiotemporal transmission risk across clinical attendee: %s percent (IQR: %s to %s)", mean_risk, q25_risk, q75_risk))

# individual transmission risk across simulations
transRisk_indv <- transRisk %>%
  group_by(patient_id) %>%
  summarize(across(c(P_N, P_N_ns), list(mean = mean, median = median,q25 = function(x) quantile(x, .25), q75 = function(x) quantile(x, .75)))) %>%
  ungroup()

transRisk_indv_pl <- transRisk_indv %>%
  mutate(patient_id = factor(patient_id)) %>%
  ggplot(aes(x = reorder(patient_id, -P_N_mean), y = P_N_mean)) +
  geom_point(shape = 4, alpha = .5) +
  geom_hline(aes(yintercept = mean_risk / 100), color = "darkred") +
  geom_hline(aes(yintercept = q25_risk / 100), color = "darkred", linetype = "dashed") +
  geom_hline(aes(yintercept = q75_risk / 100), color = "darkred", linetype = "dashed") +
  annotate("text", x = 375, y = 0.0075, label = paste0("Average across patients = ", mean_risk, "% (IQR:", q25_risk, "-", q75_risk,")"),
           size = 8 / cm(1), color = "darkred") +
  scale_y_continuous(labels = function(x) x * 100, expand = expansion(mult = c(0., .05)), breaks = c(seq(0, 0.05, 0.01), 0.004)) +
  labs(y = "Mean risk of infection (%)", x = "Clinic attendee") +
  theme_bw2() +
  theme(axis.text.x = element_blank())
save_plot(transRisk_indv_pl, pdf_file = "../results/modeling/risk-of-infection.pdf", w = aw, h = ah)

# high risk of infection plot
group_labs <- c("Small (1-5%)", "Medium (5-10%)", "High (>10%)")
transRisk_qq <- transRisk %>%
  mutate(risk_cat = base::cut(P_N * 100, breaks = c(-Inf, 1, 5, 10, Inf), labels = c("negligible risk", group_labs))) %>%
  group_by(sim, risk_cat) %>%
  summarize(n = n()) %>%
  ungroup()
transRisk_MinorVsHigh <- transRisk_qq %>%
  mutate(risk_cat = ifelse(risk_cat == "negligible risk", "low", "high")) %>%
  group_by(sim, risk_cat) %>%
  summarize(n = sum(n)) %>%
  ungroup() %>%
  group_by(sim) %>%
  mutate(share_n = n / sum(n) * 100) %>%
  ungroup() %>%
  filter(risk_cat == "high")
strMinorVsHigh <- paste0("Total risk attendees: ", 
                         round_k(mean(transRisk_MinorVsHigh$n), 0),
                         " \u00B1 ",
                         round_k(sd(transRisk_MinorVsHigh$n), 0),
                         " (",
                         round_k(mean(transRisk_MinorVsHigh$share_n), 1),
                         "% \u00B1 ",
                         round_k(sd(transRisk_MinorVsHigh$share_n), 1),
                         "%)")
transRisk_high <- transRisk_qq %>%
  dplyr::filter(risk_cat %in% group_labs) 
transRisk_qq_pl <- transRisk_high %>%
  mutate(risk_cat = factor(risk_cat, group_labs)) %>%
  ggplot(aes(x = risk_cat, y = n, fill = risk_cat)) +
  stat_boxplot(geom ='errorbar', width = .2) +
  geom_boxplot(width = .2, outlier.shape = NA) +
  annotate("text", x = .5, y = 18.5, label = strMinorVsHigh, size = 8 / cm(1), hjust = 0, vjust = 0) +
  scale_y_continuous(sec.axis = sec_axis(~ . / nrow(transRisk_indv), labels = scales::percent, name = "Proportion of attendees"),
                     expand = c(0,0), limits = c(0,20)) +
  scale_fill_manual(values = wes_palette("Darjeeling2")) +
  labs(x = "Risk of infection", y = "Number of attendees") +
  theme_bw2() +
  theme(legend.position = "none")
save_plot(transRisk_qq_pl, pdf_file = "../results/modeling/high-risk-of-infection.pdf", w = aw, h = ah)
```

### Quanta concentration

```{r}
# average quanta concentration
quantaDF_av <- sim_results %>%
  mutate(daytime = ifelse(hhmm <= "12:00", "Morning", "Afternoon"),
         daytime = factor(daytime, levels = c("Morning", "Afternoon"))) %>%
  group_by(daytime, id, location) %>%
  summarize(N = mean(N)) %>%
  ungroup()

a_ac_grid = -9
quantaDF_av_wr <- quantaDF_av %>%
  group_by(daytime) %>%
  head(nrow(waiting_room_df)) %>%
  mutate(id = as.character(id)) %>%
  ungroup() %>%
  left_join(waiting_room_df, by = c("id")) %>%
  rename(x = long, y = lat) %>%
  rotate_xy(a_ac_grid)
quantaDF_av_pa <- quantaDF_av %>%
  group_by(daytime) %>%
  tail(nrow(passage_df)) %>%
  mutate(id = as.character(1:n())) %>%
  ungroup() %>%
  left_join(passage_df, by = c("id")) %>%
  rename(x = long, y = lat) %>%
  rotate_xy(a_ac_grid)
quantaDF_av_tb <- quantaDF_av %>%
  filter(location == "tb room") %>%
  group_by(daytime) %>%
  mutate(id = as.character(1:n())) %>%
  ungroup() %>%
  left_join(tb_room_df, by = c("id")) %>%
  rename(x = long, y = lat) %>%
  rotate_xy(a_ac_grid)

qc_pl <- ggplot(mapping = aes(x = x, y = y, group = group, fill = N)) +
  geom_polygon(data = quantaDF_av_wr) +
  geom_polygon(data = quantaDF_av_pa) +
  geom_polygon(data = quantaDF_av_tb) +
  coord_fixed() +
  #scale_fill_stepsn(colours = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd")[c(1,3,5,7,9)], breaks = seq(1,6,1)*1e-6, limits = c(0,100), na.value = NA) +
  facet_wrap(~daytime, ncol = 1)
qc_pl
qc_pl <- spatial_ggpl(qc_pl)

qc_pl
save_plot(qc_pl, pdf_file = "../results/modeling/quanta-concentration.pdf", w = aw, h = 2*ah)
```

